### 2012-09-01 notes/notes.tex

\documentclass{article}

\usepackage{amsfonts}
\usepackage{amssymb}

\begin{document}

hi

\end{document}