### 2012-09-01 notes/notes.tex
changes

\documentclass{article}

\usepackage{amsfonts}
\usepackage{amssymb}

\begin{document}

\section{Computation Complexity and Intractability}

\subsection{NP-Completeness}

To prove that $\prod$ is NP-complete, we merely show that \cite{Garey:1990:CIG:574848}

\begin{enumerate}
\item $\prod \in NP and,$
\item some known NP-complete problem $\prod^\prime$ transforms to $\prod$.
\end{enumerate}

\bibliographystyle{plain}
\bibliography{ehsan}

\end{document}