### 2013-03-22 papers/coalition/fixes-extension-ideas.tex
Changes

\documentclass{article}

\usepackage{amsfonts}
\usepackage{amssymb}

\begin{document}

\section{Coalition}

\subsection{Fixes}

This domain can be modeled as a cooperative game.

Add that weights equal in expirement section.

However, unlike [1] we study...

The convexity prove has mistakes.

%\bibliographystyle{plain}
%\bibliography{ehsan}

\end{document}