### 2013-08-21 proposal/proposal/chapter3.tex

\setcounter{chapter}{2}
\chapter{Contribution and Research Activities}

%******************** START Contribution and Research Activities. ***************
In this chapter we will introduce the important notions in argumentation-based negotiation,
\section{Argumentation System and Agent Theory}\label{sec:Argumentation}