### 2013-08-22 proposal/proposal/chapter4.tex

\setcounter{chapter}{3}

\chapter{Cooperative Game Solution Concepts}

%******************** START Contribution and Research Activities. ***************
In this chapter we will introduce the important notions in argumentation-based negotiation,
\section{Argumentation System and Agent Theory}\label{sec:Argumentation}