### 2013-10-26 notes/latex.tex
changes

\documentclass{article}

\usepackage{amsfonts}
\usepackage{amssymb}

\begin{document}

\section{Latex}

Figure \ref{fig1.0} summarizes the whole proposed approach where the certainty index ($CI$) and weighted certainty index ($W\_CI$) of the moves and the whole dialogue are
measured for Type I. For Type II, different cases are considered depending on how many arguments are in use.

\begin{figure}[t]
                \begin{center}
                \includegraphics[width=15cm, height=11cm]{Figures/Approach1.eps}
                \end{center}
                \caption{General overview of the approach}\label{fig1.0}
                \label{Stack}
                \end{figure}

 \emph{(Farness Degree)}~\cite{Marey09}.

\bibliographystyle{plain}
\bibliography{ehsan}

\end{document}