### 2013-10-30 proposal/proposal/chapter2.tex
changes

\setcounter{chapter}{1}
%*******************************START Background **********************************

\chapter{Background and Relevant Literature}\label{sec:MAS}

    \section{Community of Web Services}\label{sec:CommunityWS}

        \subsection{Web Services}\label{sec:WS}

        \subsection{Definition}\label{sec:CWSDefinition}

        \subsection{Architecture}\label{sec:CWSArchitecture}

        \subsection{Operations}\label{sec:CWSOperations}

            \subsubsection{Community Development}\label{sec:CWSCommunityDev}

            \subsubsection{Web Service Attraction and Retention}\label{sec:CWSAttraction}

            \subsubsection{Web Service Selection}\label{sec:CWSSelection}

    \subsection{Related Work}\label{sec:CWSRelatedWork}

%*******************************End Background **********************************