### 2013-11-03 proposal/proposal/chapter2.tex

\setcounter{chapter}{1}
%*******************************START Background **********************************

\chapter{Background and Relevant Literature}\label{sec:MAS}

    \section{Community of Web Services}\label{sec:CommunityWS}

        \subsection{Web Services}\label{sec:BRWS}

        \subsection{Definition}\label{sec:CWSDefinition}

        \subsection{Architecture}\label{sec:CWSArchitecture}

        \subsection{Operations}\label{sec:CWSOperations}

            \subsubsection{Community Development}\label{sec:CWSCommunityDev}

            \subsubsection{Web Service Attraction and Retention}\label{sec:CWSAttraction}

            \subsubsection{Web Service Selection}\label{sec:CWSSelection}

    \section{Cooperative Game Theory in Multiagent Systems}\label{sec:CGTMS}

        \subsection{Web Services}\label{sec:BRWS}

        \subsection{Definition}\label{sec:CWSDefinition}

        \subsection{Architecture}\label{sec:CWSArchitecture}

    \section{Related Work}\label{sec:BRRelatedWork}

%*******************************End Background **********************************