### 2013-11-05 proposal/proposal/Abstract.tex
Changes

\begin{center}
{\LARGE\textbf{Abstract}}
\end{center}

Web services are loosely-coupled business applications willing to
cooperate in distributed settings within different groups called
communities. Communities aim to provide better visibility,
efficiency, market share and total payoff. There are a number of
proposed mechanisms and models on aggregating web services and
making them cooperate within their communities. However, forming
optimal and stable communities as coalitions to maximize
individual and group efficiency and income has not been addressed
yet.

In this proposal, we propose an efficient coalition formation
mechanism using cooperative game-theoretic techniques. We propose
a mechanism for community membership requests and selections of
web services in the scenarios where there is interaction between one community and many web services
and scenarios where web services can join multiple established communities. The ultimate objective is to
develop a mechanism for web services to form stable groups
allowing them to maximize their efficiency and generate
near-optimal (welfare-maximizing) communities. The theoretical and
extensive simulation results show that our algorithms provide web services
and community owners, in real-world like environments, with applicable and near-optimal decision
making mechanisms.

In future work, we need to apply different cooperative game techniques such as nucleus, 
kernel and bargaining solution concepts and theoretically analyse solution concepts, 
based on functions representing payoff and quality metrics the web service communities. 
We also aim to apply reinforcement learning techniques for individual web service 
strategic decision making process.