### 2013-11-07 proposal/proposal/chapter4.tex

\setcounter{chapter}{3}

\chapter{Cooperative Game Solution Concepts}

This chapter was merged with chapter 2