### 2013-12-11 proposal/presentation/ekpresentation/ehsan_presentation.tex
changes

\documentclass{beamer}
\usepackage[latin1]{inputenc,colortbl}
\usepackage{epsfig}
\usetheme{Frankfurt}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}[page number]
%\newtheorem{definition}{Definition}
\title[Cooperative Decision Making for Communities of Autonomous Web Services]{Cooperative Decision Making for Communities of Autonomous Web Services}
\author{Ehsan Khosrowshahi Asl\\ \vspace{0.2cm} Supervised by: \\Dr. Jamal Bentahar \\Dr. Hadi Otrok}
\institute{Department of Computer Science and Software Engineering\\Concordia University}
\date{December 17, 2013}
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame1 title page %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\titlepage
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame2 outline page %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Presentation Outline}
    \begin{itemize}
     	\itemsep=.5cm
    	\item {\bf Introduction}
    	\item Background and Literature Review
    	\item Proposed Research
        \item Contributions and Research Activities
    	\item Time Line
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame3 introduction %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%\section{Motivation}
\subsection{Motivation}
    \begin{frame}{Motivation}
    \begin{itemize}
        \item High reliance on web services, created high competition and great business opportunities for online service developers
        \item Community goals:
        \begin{itemize}
            \item \textbf{User satisfaction, Service quality, high availability and responsiveness}.
            \item \textbf{Web Service Discovery over Internet}.
        \end{itemize}
    \end{itemize}
    \begin{figure}[htbp]
        \centering
        %\includegraphics[width=12cm, height=8cm]{figures/figure1.eps}
        \includegraphics[width=1.0 \columnwidth]{figures/figure7.png}
        %%\includegraphics[scale=0.5]{figure1}
        %\caption{The NetBill payment protocol} \label{figure7}
    \end{figure}
    \end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Problems and Research Questions}

\begin{itemize}
    \item \textbf{How can we model the
        community of agent-based services in order to maximize the utility
        of involved users, web services and community organizers?}.
        \begin{itemize}
            \item Most of the recent work on communities of
            services are either user-centric and focus on user satisfaction
            or system-centric and focus on the whole system throughput, performance and utilization.
        \end{itemize}

    \item \textbf{How can we model fair and stable communities as coalitions
        of agent-based web services?}.
        \begin{itemize}
            \item All parties involved are self-interest agents and have the authority to deviate leave the community if they can improve their income in some other ways.
        \end{itemize}

\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Continue: Problems and Research Questions}
 \begin{itemize}
   \item \textbf{How can we model and analyze the cooperation
        among the community members in realistic, applicable and practical
        settings?}
        \begin{itemize}
            \item Modeling the preference and valuation functions considering the worth of all subsets of players in realistic cooperative scenarios
            \item Solution concepts complexity issues
        \end{itemize}

\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    \begin{frame}{Multi-Agent Systems MASs}
%        \begin{itemize}
%            \itemsep=.35cm
%        	\item Multi-Agent System (MAS).
%            %\item Autonomous agents.
%        	\item Knowledge in MASs.
%            \item Social commitments in MASs.
%        \end{itemize}
%\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background}
\begin{frame}{Presentation Outline}
    \begin{itemize}
     	\itemsep=.5cm
    	\item Introduction
    	\item {\bf Background and Literature Review}
    	\item Proposed Research
        \item Contributions and Research Activities
    	\item Time Line
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame9 Background and Literature Review %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Communities of Web Services}
    \begin{frame}{Community of Web Services}
        Communities of Web Services
        \begin{itemize}
            \item Ease and improve the process of Web services discovery in an open environment like the Internet
        	\item Improve service quality, availability and responsiveness
        \end{itemize}

        \begin{figure}[htbp]
            \centering
            %\includegraphics[width=12cm, height=8cm]{figures/figure1.eps}
            \includegraphics[width=1.0 \columnwidth]{figures/wscommunity2.png}
            %%\includegraphics[scale=0.5]{figure1}
            %\caption{The NetBill payment protocol} \label{figure7}
        \end{figure}

    \end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame10 Background and Literature Review %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Social commitments in MASs}
    \begin{frame}{Communities of Web Services}
        Communities of Web Services
        \begin{itemize}
            \itemsep=.35cm
        	\item \textbf{Communities of web services operations}
            \begin{itemize}
              \item Community Development
              \item Web Services Attraction and Retention
              \item Web Service Selection
            \end{itemize}                      	      	
        \end{itemize}

        \begin{figure}[htbp]
            \centering
            %\includegraphics[width=12cm, height=8cm]{figures/figure1.eps}
            \includegraphics[width=1.0 \columnwidth]{figures/wscommunity2.png}
            %%\includegraphics[scale=0.5]{figure1}
            %\caption{The NetBill payment protocol} \label{figure7}
        \end{figure}
    \end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame11 Background and Literature Review %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cooperative Game Theory}

    \begin{frame}{Cooperative Game Theory}
      \begin{itemize}
         \item Cooperative games are a branch of game theory that models cooperation or collaboration between agents.
         \begin{itemize}
            \item A set of tasks needs to be performed, they require different expertise.
            \item Agents do not have enough resource on their own to perform a task.
            \item Find complementary agents to perform the tasks.
            \begin{itemize}
                \item robots have the ability to move objects in a plant, but multiple robots are required to move a heavy box.
                \item transportation domain: agents are trucks, trains, airplanes, ships... a task is a good to be transported.
            \end{itemize}
        \end{itemize}
        \item \textbf{issues:}
            \begin{itemize}
                \item What coalition to form?
                \item How to reward each each member when a task is completed?
            \end{itemize}
      \end{itemize}
    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame14 Background and Literature Review %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{The Classic Problem}
    We have a population $N$ of $n$ agents.
    \begin{definition} [Coalition]~\\
       A coalition $C$ is a set of agents: $C \in 2^N$.
        \begin{itemize}
            \item $N$ is the set of all agents (or players)
            \item $v:2^N \rightarrow R$ is the \emph{valuation function}. For $C \subseteq N$, $v(C)$ is the value obtained by the coalition $C$
        \end{itemize}
    \end{definition}

    \begin{itemize}
        \item \textbf{problem:} a game $(N,v)$, and we assume all agents in $N$ want to cooperate.
        \item \textbf{solution:} a payoff distribution $x \in R^n$ that provides a value to individual agents.

        \begin{itemize}
            \item What are the interesting properties that $x$ should satisfy?
            \item How to determine the payoff vector $x$?
        \end{itemize}

    \end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{An Example}

    \begin{center}
        $N = \{1,2,3\}$ \\
        $v(\{1\}) = 0, v(\{2\}) = 0, v(\{3\}) = 0$ \\
        $v(\{1,2\}) = 90$ \\
        $v(\{1,3\}) = 80$ \\
        $v(\{2,3\}) = 70$ \\
        $v(\{1,2,3\}) = 105$ \\
    \end{center}

    What should we do?
    \begin{itemize}
        \item form $\{1,2,3\}$ and share equally $(35,35,35)$?
        \item 3 can say to 1 ``let's form $\{1,3\}$ and share (40,0,40)''
        \item 2 can say to 1 ``let's form $\{1,2\}$ and share (45,45,0)''
        \item 3 can say to 2 ``OK, let's form $\{2,3\}$ and share (0,46,24)''
        \item 1 can say to 2 and 3, ``fine! $\{1,2,3\}$ and (33,47,25)''
        \item ...what is a good solution?
    \end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame15 outline page

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Transferable and Non Transferable Utility Games}
    \begin{itemize}
        \item \textbf{Games with Transferable Utility (TU games)}
        \begin{itemize}
            \item Utility is worth the same for all agents
            \item Utility can be {\color{red} compared} or {\color{red} transfered} between agents.
            \item Irrespective of the division of the coalitional payoff, members of the coalition enjoy the same total utility.
        \end{itemize}

        \begin{definition} [Valuation or Characteristic function]\label{dfn:valuationfunction}
            A \emph{valuation function v} associates a real number $v(C)$ to any subset $C \subseteq N$, i.e., $v:2^N \rightarrow R$ \\
            A {\color{blue}\emph{TU game}} is a pair $(N,v)$ where $N$ is a set of agents and where $v$ is a valuation function.
        \end{definition}

        \item \textbf{Games with Non Transferable Utility (NTU games)} \\
            Agents have different preferences over coalitions and rewards. Non-monetary rewards, i.e. item allocation type of problems where items have different values for different players.
    \end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Some properties of valuation functions}
    $\forall C_1,C_2 \subseteq N | C_1 \bigcap C_2 = \emptyset, i \in N, i \notin C_1$
    \begin{itemize}
        \item {\color{red} additive:} $v(C_1 \bigcup C_2) = v(C_1) + v(C_2)$
        \item {\color{red} super additive:} $v(C_1 \bigcup C_2) \geq v(C_1) + v(C_2)$ This is satisfied is many applications, or bigger coalitions will not form.
        \item {\color{red} weekly super additive:} $v(C_1 \bigcup \{i\}) \geq v(C_1) + v(\{i\})$
        \item {\color{red} subadditive:} $v(C_1 \bigcup C_2) \leq v(C_1) + v(C_2)$
    \end{itemize}

    $\forall C_1,C_2 \subseteq N$
    \begin{itemize}
        \item {\color{red} convex:} $v(C_1) + v(C_2) \leq v(C_1 \bigcup C_2) + v(C_1 \bigcap C_2)$ Convexity has important properties in cooperative game theory solution concepts.
    \end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Solution properties}
    Let $x \in R^n$ be a solution of the TU game $(N,v)$
    \begin{itemize}
        \item {\color{red} Feasible solution:} $\sum_{i \in N} x(i) \leq v(N)$
        \item {\color{red} Efficiency:} $x(N) = v(N)$
        \begin{itemize}
            \item the payoff distribution is an allocation of the entire worth of the grand coalition to all agents
        \end{itemize}
        \item {\color{red} Individual rationality:} $\forall i \in N, x(i) \geq v(\{i\})$
        \begin{itemize}
            \item player obtains at least its self-value of payoff
        \end{itemize}
        \item {\color{red} Group rationality:} $\forall C \subseteq N, \sum_{i \in N} x(i) \geq v(C)$
    \end{itemize}

    \vspace{0.2cm}

    An {\color{red} imputation} is a payoff distribution $x$ that is efficient and individual rational.

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Stability}
    \begin{itemize}
        \item We all want to work together and get $v(N)$, but we all have different views about how to share the fruits of our work. We can use the values of other coalitions as arguments in favor of a distribution.
        \item A condition for a coalition to form: 
            {\color{red}all} agents prefer to be in it. i.e., none of the participants wishes she were in a different coalition or by herself {\color{red} $\Rightarrow Stability$ }
        \item The {\color{red} core} is a stability concepts for which no agents prefer to deviate to form a different coalition.
        \item Is the coalition stable $\Leftrightarrow$ Is the core non-empty
    \end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Solution Concepts: Core}
    CTLC logic is an extension of CTL logic with modalities for reasoning about \textbf{commitments and their fulfillment}.
    \begin{definition} [Imputation]\label{dfn:imputationbig}
        An {\color{red}imputation} is a payoff distribution $x$ that is efficient and individually rational.
        \begin{itemize}
            \item $\sum_{i \in N} x(i) = v(N)$ (efficiency)
            \item for all $i \in N, x_i \geq v(\{i\})$ (individual rationality)
        \end{itemize}        
    \end{definition}

    \begin{definition} [Group rationality]\label{dfn:grouprationality}
        $\forall C \subseteq N, \sum_{i \in N} x(i) \geq v(C)$
    \end{definition}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} {Solution Concepts: Core}
    The core relates to the stability of the grand coalition: \\ No group of agents has any incentive to change coalition.
    \begin{definition}[$Core$ of a game $(N,v)$]\label{dfn:core}
        Let $(N,v)$ be a cooperative game, and assume they form the coalition $N$. The core of $(N,v)$ is the set:
        \vspace{0.1cm}
        \begin{center}
            $Core(N,v) = \{x \in R^n | x$ is a group rational imputation$\}$        
        \end{center}
        Equivalently,
        \vspace{0.1cm}
        \begin{center}
            $Core(N,v) = \{x \in R^n | x(N) \leq v(N) \wedge x(C) \geq v(C) \forall C \subseteq N\}$
        \end{center}
    \end{definition}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} {Example: Core}

    \begin{center}
      $N = \{1,2\}$ \\
      $v(\{1\}) = 5, v(\{2\}) = 5$ \\
      $v(\{1,2\}) = 20$ \\      
    \end{center}

    $Core(N,n) = \{(x_1,x_2) \in R^2 | x_1 \geq 5, x_2 \geq 5, x_1 + x_2 = 20\}$

    \begin{figure}[htbp]
        \centering
        %\includegraphics[width=12cm, height=8cm]{figures/figure1.eps}
        \includegraphics[width=0.3 \columnwidth]{figures/coreex1.png}
        %%\includegraphics[scale=0.5]{figure1}
        %\caption{The NetBill payment protocol} \label{figure7}
    \end{figure}

    The core may not be fair: the core only considers stability.

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Proposed Research}
\begin{frame}{Presentation Outline}
    \begin{itemize}
     	\itemsep=.5cm
    	\item Introduction
    	\item Background and Literature Review
    	\item {\bf Proposed Research}
        \item Contribution and Research Activities
    	\item Time Line
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame16 Proposed Research %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Combining Logics}

\begin{frame}{A schematic view of our approach}

\begin{figure}[htbp]
\centering
%\includegraphics[width=12cm, height=8cm]{figures/figure1.eps}
\includegraphics[width=1.05 \columnwidth]{figures/figure9.png}
%%\includegraphics[scale=0.5]{figure1}
%\caption{Our approach} %\label{figure9}
\end{figure}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}{Combining Logics}
Given \textbf{two logics} $L_1$ and $L_2$, their \textbf{languages} $\mathcal{L}_1$ and $\mathcal{L}_2$, and their
\textbf{axiomatic systems} $\mathcal{X}_1$ and $\mathcal{X}_2$, the logic
$L_1 \otimes L_2$ is the smallest logic with the following
characteristics:
\begin{itemize}
  \item The language $\mathcal{L}_{L_1\otimes L_2}$ is the union of $\mathcal{L}_{L_1}$ and $\mathcal{L}_{L_2}$.
  \item The logic $L_1 \otimes L_2$ is axiomatised by the set of axioms $\mathcal{X}_1 \cup \mathcal{X}_2$.
\end{itemize}
If the logics $L_1$ and $L_2$ are interpreted using Kripke frames
$F_1 = (S, R_1^1, \dots, R_n^1)$ and $F_2 = (S, R_1^2, \dots
,R_m^2 )$ where $R^1_{x/x\in[1,n]}$ and $R^2_{y/y\in[1,m]}$ are
\textbf{accessibility relations}, then the semantics for $L_1 \otimes L_2$
can be defined in the Kripke frame $F =(S, R_1^1, \dots , R_n^1,
R_1^2, \dots ,R_m^2 )$ obtained by the \textit{independent join} of the two
frames $F_1$ and $F_2$.
    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame17 Proposed Research
\begin{frame} {Model of CTLKC}
\begin{definition} [Model of CTLKC]\label{Model of CTLKC}
A model $\mathcal{M}=(S,I,R_t,\{\approx_i|i\in \mathcal{ A}\},\{\sim _{i \rightarrow j}|{(i,j)\in \mathcal{A}^2}\},\mathcal{V})$ that belongs to the set of all models $\mathbb{M}$ is a tuple, where:
\begin {itemize}
\item $S, I, R_t$ and $\mathcal{V}$ are the same as in Definition (Model of CTLK);
\item For each agent $i \in \mathcal{A}$, $ \approx _i  \subseteq S \times S
$ is the epistemic accessibility relation defined by $ s \approx_i
s' $ iff $ l_i(s) = l_i(s') $;
\item For each pair $(i, j) \in \mathcal{A}^2 $, $\sim _{i \rightarrow j} \subseteq S \times S $ is the social accessibility relation defined by $ s \sim _{i \rightarrow j} s' $ iff
     1) $ l_i(s) = l_i(s') $ and
     2) $ Var_i \cap Var_j \neq \emptyset $ such that  $ \forall x \in Var_i \cap Var_j $ we have $ l_i^x(s) = l_j^x(s') $ and $ \forall y \in Var_j - Var_i $ we have $ l_j^y(s) = l_j^y(s')$;
\item $ \mathcal{V} : S \rightarrow 2 ^{\Phi_p} $ is a valuation function where $\Phi_p $ is a set of atomic propositions.
\end {itemize}
\label {dfn: Model of CTLKC}
\end{definition}
 \end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} {Syntax of CTLKC}

\begin{definition}[Syntax of CTLKC]~\\
$ ~~~~~~~~~~~~~\varphi ::= p \mid \neg \varphi \mid \varphi \vee
\varphi \mid EX \varphi  \mid E ( \varphi U \varphi) \mid EG
\varphi\mid K_i \varphi \mid C_{i\rightarrow j} \varphi \mid Fu
(C_{i\rightarrow j}\varphi)$. \label {dfn: Syntax of CTLKC}
\end{definition}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame18 Proposed Research %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Selection Criteria}
We propose the following criteria based on \textbf{common intuitions}:
\begin{itemize}
%\begin{enumerate}
  \item It should not be the case that an agent commits about everything he knows to others.
  \item It should not be the case that an agent commits about everything known by others.
  \item An agent should know about his own commitment.
  \item An agent should be aware of any commitment directed towards him.
  \item An agent should know the content of his fulfilled commitment.
  \item An agent should know the content of any fulfilled commitment directed towards him.

\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Continue: Selection Criteria}
\begin{itemize}
 \item An agent should know about the fulfillment of his commitment.
 \item An agent should know about the fulfillment of any commitment directed towards him.
  \item A sincere agent should know (i.e., believe) the content of his own commitment.
  \item A sincere agent should fulfill his commitment.
  \item A liar agent could not know (i.e., believe) the content of his commitment.
  \item If an agent knows about his own commitment, then this commitment should exist, which means an agent should not know about a non-existing commitment.
\end{itemize}
%\end{enumerate}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Running Example: NetBill Payment Protocol}
%
\begin{figure}[htbp]
\centering
%\includegraphics[width=12cm, height=8cm]{figures/figure1.eps}
\includegraphics[width=.75 \columnwidth]{figures/figure1.png}
%%\includegraphics[scale=0.5]{figure1}
%\caption{The NetBill payment protocol} \label{figure1}
\end{figure}
\begin{itemize}
\item $C_{Cus \rightarrow Mer}$ $sendPayment$.
 \item $C_{Mer \rightarrow Cus}$ $deliverGoods$.
 \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame19 Proposed Research
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Desiderata of Paradoxes}

\begin{frame}{Desiderata of Paradoxes}
Paradoxes are represented as reasoning rules that are: 1) not reasonable in
open MAS, but they are valid in CTLKC; or 2) desirable for the
interaction between knowledge and commitments, but they can get
violated in CTLKC because they are not valid.
 \begin{enumerate}

\vspace{0.2cm} \item [P1.]\textbf{[Committing everything known to
others]}

$K_i \varphi \Rightarrow C_{i \rightarrow j}\varphi, $ for all $j
\in A$ where $i \neq j$.

\textbf{ Meaning:} An agent (the debtor) commits towards all other
agents (the creditors) about what he knows.\\
\textbf{Example 1:}
$K_{Mer}$ $deliverGoods \Rightarrow$ $C_{Mer \rightarrow
Cus}$ $deliverGoods$.

\vspace{0.2cm} \item [P2.]\textbf{[Committing everything known by
others]}

 $ K_i K_j \varphi \Rightarrow C_{i \rightarrow j} \varphi$ such that $ i \neq j$.

\textbf{Meaning:} An agent (the debtor) commits toward another
agent (the creditor) to bring about what he (i.e., the debtor)
knows about the creditor's knowledge.\\
\textbf{Example 2:}
$K_{Cus} K_{Mer}$ $deliverGoods \Rightarrow$ $C_{Cus \rightarrow Mer}$
$deliverGoods$.

\end{enumerate}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame20 Proposed Research %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Continue: Desiderata of Paradoxes}
 \begin{enumerate}
\vspace{0.2cm} \item [P3.][\textbf{Committing everything known
from others}]
$ K_i K_j \varphi \Rightarrow C_{j \rightarrow i} \varphi$ such that $ i \neq j$.

\textbf{Meaning:} An agent (the debtor) commits toward another
agent (the creditor) to bring about what the creditor knows about
the debtor's knowledge.\\
\textbf{Example 3:}\\
$K_{Cus} K_{Mer}$ $deliverGoods \Rightarrow$ $C_{Mer \rightarrow Cus}$
$deliverGoods$.

\vspace{0.2cm} \item [P4.][\textbf{Knowing about his own
commitment}]

$C_{i\rightarrow j} \varphi \Rightarrow K_i (C_{i\rightarrow j}
\varphi) $ where $i \neq j$.

\textbf{Meaning:} An agent knows about his commitment.

\textbf{Example 4:}\\
$C_{Cus \rightarrow Mer}$ $sendPayment $ $\Rightarrow K_{Cus}$ $(C_{Cus \rightarrow Mer}$ $sendPayment)$.
\end{enumerate}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Continue: Desiderata of Paradoxes}

\begin{figure}[htbp]
    \begin{center}
%\includegraphics[width=10cm, height=5cm]{figures/figure2.eps}
    \includegraphics[width=.75 \columnwidth]{figures/figure2.png}
   % \caption{Model 1} \label{figure2}
    \end{center}
    \end{figure}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Continue: Desiderata of Paradoxes}
\begin{enumerate}
\vspace{0.2cm} \item [P5.][\textbf{Knowing the content of his own
fulfilled commitment}]

$Fu (C _{i\rightarrow j} \varphi) \Rightarrow K _i \varphi$ where
$i \neq j$.

\textbf{Meaning:} An agent knows the content of his fulfilled
commitment.\\
\textbf{Example 5:}\\
$Fu (C _{Cus\rightarrow Mer} sendPayment) \Rightarrow K_{Cus}$ $sendPayment$.
\begin{figure}[htbp]
    \begin{center}
%\includegraphics[width=10cm, height=5cm]{figures/figure2.eps}
    \includegraphics[width=.75 \columnwidth]{figures/figure3.png}
   % \caption{Model 2} \label{figure3}
    \end{center}
    \end{figure}

\end{enumerate}     	
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame23 Proposed Research %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

 \begin{frame}{Continue: Desiderata of Paradoxes}
 \begin{enumerate}

\vspace{0.2cm} \item [P6.][\textbf{Knowing the content of the
other's fulfilled commitment}]

$Fu (C _{i\rightarrow j} \varphi) \Rightarrow K_j \varphi$ where
$i \neq j$.

\textbf{Meaning:} The creditor knows the content of the debtor's
fulfilled commitment.\\
\textbf{Example 6:}\\
$ Fu (C_{cust \rightarrow Mer}$ $sendPayment)$ $\Rightarrow $ $k_{Mer} sendPayment$.
	\begin{figure}[htbp]
\begin{center}
%\includegraphics[width=12cm, height=5cm]{figures/figure4.eps}
\includegraphics[width=.70 \columnwidth]{figures/figure4.png}
%\caption{Model 3} \label{figure4}
\end{center}
\end{figure}
\end{enumerate}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame24 Proposed Research %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Continue: Desiderata of Paradoxes}
\begin{enumerate}
    \vspace{0.2cm} \item [P7.][\textbf{Knowing the commitment of the
other}]

$C_{i\rightarrow j} \varphi \Rightarrow AF K_j (C_{i\rightarrow j}
\varphi) $ where $i \neq j$.

\textbf{Meaning:} An agent should be aware of any commitments
directed towards him.\\
\textbf{Example 7:}\\
$C_{Mer \rightarrow Cus}$ $deliverGoods$ $\Rightarrow $ $AF K_{Cus}(C_{Mer \rightarrow Cus}$
$deliverGoods$).
\begin{figure}[htbp]
\centering
\includegraphics[width=.50 \columnwidth]{figures/figure5.png}
%\includegraphics[width=12cm, height=7cm]{figures/figure5.eps}
%\includegraphics[scale=0.5]{figure4.eps}
%\caption{Model 4} \label{figure5}
\end{figure}

\end{enumerate}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Continue: Desiderata of Paradoxes}
\begin{enumerate}
    \vspace{0.2cm} \item [P8.][\textbf{Knowing the fulfillment of his
own commitment}]

$Fu(C _{i\rightarrow j} \varphi) \Rightarrow K_i Fu(C
_{i\rightarrow j} \varphi)$ where $i \neq j$.

\textbf{Meaning:} The agent knows that he fulfills his commitment.\\
\textbf{Example 8:}\\
$Fu(C_{Cus\rightarrow Mer} sendPayment)\Rightarrow K_{Cus}Fu(C_{Cus \rightarrow Mer} sendPayment)$.

\vspace{0.2cm} \item [P9.][\textbf{Knowing the fulfillment of the
debtor's commitment}]

$Fu(C _{i\rightarrow j} \varphi) \Rightarrow K_j Fu(C_{i\rightarrow j} \varphi)$ where $i \neq j$.

\textbf{Meaning:} The creditor knows that the debtor fulfills his
commitment.\\
\textbf{Example 9:}\\
$Fu(C_{Cus \rightarrow Mer} sendPayment)\Rightarrow K_{Mer} Fu(C_{Cus \rightarrow Mer} sendPayment)$.
\end{enumerate}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame25 outline page %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{CTLKC properties}
\begin{frame}{CTLKC properties}
In addition to the properties of knowledge (in CTLK) and commitments (in CTLC), those properties hold in CTLKC:
\begin{enumerate}

\vspace{0.2cm} \item [Prop 1.] $C_{i \rightarrow j}\varphi
\Rightarrow \varphi$ is satisfiable but not valid.
\vspace{0.2cm} \item [Prop 2.] $C_{i \rightarrow j}\varphi
\Rightarrow K_i \varphi$ is satisfiable but not valid.
\vspace{0.2cm} \item [Prop 3.] $(C_{i \rightarrow j}\varphi \wedge
K_i (\varphi \Rightarrow \psi)) \Rightarrow C_{i \rightarrow j}
\psi$ is valid.
\vspace{0.2cm} \item [Prop 4.] $(C_{i \rightarrow j}(\varphi
\Rightarrow \psi) \wedge K_i \varphi) \Rightarrow C_{i \rightarrow
j} \psi$ is valid.
\vspace{0.2cm} \item [Prop 5.] $(C_{i \rightarrow j}\varphi \wedge
K_j (\varphi \Rightarrow \psi)) \Rightarrow C_{i \rightarrow j}
\psi$ is satisfiable but not valid.
\vspace{0.2cm} \item [Prop 6.] $(C_{i \rightarrow j}(\varphi
\Rightarrow \psi) \wedge K_j \varphi) \Rightarrow C_{i \rightarrow
j} \psi$ is satisfiable but not valid

\end{enumerate}
\end{frame}

\section{Research Activities}
\begin{frame}{Contribution and Research Activities}
    \begin{itemize}
     	\itemsep=.5cm
    	\item Introduction
    	\item Background and Literature Review
    	\item Proposed Research
        \item {\bf Contribution and Research Activities}
    	\item Time Line
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame26 Contribution and Research Activities %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Contributions}
    \begin{frame}{Contributions}

    \begin{itemize}
     \itemsep=.35cm
     \item A new combined temporal logic, called CTLKC, that combines both CTLK and CTLC logics is introduced.
     \item This logic served as a language to express a set of rules that are used to reason about the interaction between knowledge and social commitments.
     \item Selection criteria are proposed to capture such reasoning rules.
     \item Nine paradoxes are recognized in CTLKC logic that should be addressed in any consistent logic combining knowledge and commitments.
     \item Six properties are identified and should be met in any logic combining knowledge and commitments.
        \end{itemize}

    \end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame27 Contribution and Research Activities

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame30 outline page
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Milestones}
\begin{frame}{Milestones}
    \begin{itemize}
     	\itemsep=.5cm
    	\item Introduction
    	\item Background and Literature Review
    	\item Proposed Research
        \item Contribution and Research Activities
    	\item {\bf Time Line}
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Research Activities Goal}
    \begin{frame}{Research Activities Goal}
    We aim to cover all possible relations between knowledge and commitments as single and group interactions by:

        \begin{itemize}
              \item Defining a new consistent logic that combines knowledge and commitments and solves the nine paradoxes while satisfying the six properties.
              \item Addressing the model checking problem of the new logic from both algorithmic and complexity perspectives.
              \item Implementing the new model checking algorithms on top of MCMAS+.
              \item Identifying a new semantics for group committing.
              \item Adding the group knowledge to the framework.
        \end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Time Line}
    \begin{frame}{Time Line}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=.65 \columnwidth]{figures/figure6.png}
%\includegraphics[width=16cm, height=21cm]{figures/figure6.eps}
%\caption{Schedule time of research} \label{figure6}
\end{center}
\end{figure}
    \end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame27 Milestones %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Publications}

1- Submitted a journal paper entitled ``\textbf{Interaction between knowledge and Social Commitments in MASs}", Applied Intelligence Journal, Springer.\\

%\centering
%\begin{itemize}
%\item \textbf{Questions Please !}
%\end{itemize}

\end{frame}

\end{document}