### 2014-01-01 papers/coalition-journal-R1/response_letterr.tex
changes

%\documentclass[preprint,authoryear,12pt]{elsarticle}
\documentclass[times, 12pt,a4paper]{article}
\usepackage{graphicx,amssymb,amsthm,stmaryrd}%,booktabs}
\usepackage{times}
\usepackage{amsmath,url,algorithm,algorithmic,mathrsfs,makeidx}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{epsfig}
\usepackage{latex8}

%journal{Journal of Knowledge-Based Systems}

\begin{document}
%\begin{frontmatter}

\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{example}{Example}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary} {Corollary}
\title{\underline{Response Letter}\\\vspace{0.5cm}Efficient Community Formation for Web Services}
\author{Ehsan Khosrowshahi Asl\\
e\_khosr@encs.concordia.ca\\
\and
Hadi Otrok\\
hadi.otrok@kustar.ac.ae \\
\and
Jamal Bentahar\\
bentahar@ciise.concordia.ca \\
\and
Rabeb Mizouni\\
rabeb.mizouni@kustar.ac.ae \\
}

\maketitle

First, we would like to thank and express our appreciation to the
reviewers for taking the time to carefully review our manuscript,
and for their insightful, valuable, and very useful comments. We
would like to thank the editor as well for his valuable comments
and suggestions for improvements. We carefully considered the
comments and implemented all of them in the revised version. In
this letter, we will explain, point by point, how the issues
raised in the reviews are addressed and answered.

%\newpage

%%%%%%%%%%%%%%%%%%%%%%%%% REVIEWER 1 %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{center}
  \textbf{Reviewer 1}
\end{center}

\vspace{0.4cm}\textbf{\underline{1. Reviewer}:}
You should relate V(S) to PO(C) in equations 5 and 6.

\vspace{0.2cm}\textbf{\underline{Answer 1}:
We modified the section and added the connection at the end of paragraph,
however we had already connected them when we were describing the algorithm
of our scenario in section 4.1}

\vspace{0.4cm}\textbf{\underline{2. Reviewer}:}
please add some definitions such as the one for Grand coalition
to make the paper self-contained.

\vspace{0.2cm}\textbf{\underline{Answer 2}:
Great suggestion, we did add more definitions for second scneraio too.}

\vspace{0.4cm}\textbf{\underline{3. Reviewer}:}
Why you didn't leverage the foundations of repeated games instead of
coalitional games? Is there any strong reason?.

\vspace{0.2cm}\textbf{\underline{Answer 3}:
The communities are models as cooperative groups and that is why cooperative game thoery methods provide us with great decision mechanisims for forming groups and coalitions which would benefit all the agents involved altogether. In repeated games, there should be a repeatable game between agents one to one, where decision cannot be optimal for all the agents involved. We belieave repeated game is appropriate in cases where there is a lot of competing one-to-one interaction between agents are happening, and actually we are working on repeated games and specificly repeated game with learning techniques such as q-learning to help distribute the ``task distribution job'' among all the web services. Where web services are having different confidence states, and based on their action history and state they are in, web services can directly opt for getting and performing tasks alone or cooperating which other agents.
}

\vspace{0.4cm}\textbf{\underline{4. Reviewer}:}
What do you mean by "the algorithm gets interrupted": is it based on
a logical condition? if yes, what is it?

\vspace{0.2cm}\textbf{\underline{Answer 4}:
It means if there is time constraint, the algorithm can get interrupted any time, and yet have depth(n-1) results. Each depth will have $O(n^2)$ cases to check, so for example if in 5 second it reaches depth 10, and we interrupt the algorithm since we cannot wait any loner,  we at least have checked for all subsets of depth 9}

\vspace{0.4cm}\textbf{\underline{5. Reviewer}:}
Please provide equations using lambda and epsilon and relate them
to your model that uses Rc.

\vspace{0.2cm}\textbf{\underline{Answer 5}:
We added the new stability($Core$ condition) in case of having taxtation.}

\vspace{0.4cm}\textbf{\underline{5. Reviewer}:}
For the epslion core method: the value of epsilon should be
varied to verify its impact in the experiments section.

\vspace{0.2cm}\textbf{\underline{Answer 5}:
That is exactly what we are doing, we changes the explanation to make what we did more clear.}

\vspace{0.4cm}\textbf{\underline{6. Reviewer}:}
The experiments are missing a comparison with [18] is missing.

\vspace{0.2cm}\textbf{\underline{Answer 6}:
The expirement which was illustrated in figure[9] was comparision of our work with [18], for all the scenarios which we could make it comparable and fair with. We have called their method ``High Availability Coalition Method''}

\vspace{0.4cm}\textbf{\underline{7. Reviewer}:}
Scalability evaluation is missing: what is the impact of increasing the number of services/communities on the system efficiency?.

\vspace{0.2cm}\textbf{\underline{Answer 7}:
As mentioned in the paper, our algorithms run in $O(n)$ and $O(n^2)$ time as number of web services with a community (n) grows. Also with the real web service data provided in http://www.uoguelph.ca/\textasciitilde{}qmahmoud/qws/ the largest communities in our expirements, did not grow larger than 60 web services. However in cases where $R_C$ is very large, communities can grow
large. However there will be no problem community management, and task distrobution algorithms as the number of web services grows, I could set $R_C$ very large in simulation but since we do not evaluate run time of comunity formation and membership strategies decisions, it will not affect our results.
}

%%%%%%%%%%%%%%%%%%%%%%%%% REVIEWER 2 %%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{2cm}

\newpage

\begin{center}
  \textbf{Reviewer 2}
\end{center}

\vspace{0.4cm}\textbf{\underline{1. Reviewer}:}  Though there’s a referral to [8] it’d be helpful to explain more the concept of “Community coordinators” web services, somewhere in the early parts of the paper. How’s it different from other (common) web services in a given community.

``community master'' and ``community coordinator'' – are these different?

\vspace{0.2cm}\textbf{\underline{Answer 1}: We changed the way how we described our architecture, we dropped the concept of community master web service and we defined a separate entity as community coordinator.}

\vspace{0.4cm} \textbf{\underline{2. Reviewer}}: p.10 ``our community receives 130 tasks from users'' what are the types of tasks. Are these simulated or real functions that need to performed by web services?

\vspace{0.2cm}\textbf{\underline{Answer 2}: The tasks are randomly generated, and since communities are composed of web services performing same type of operations, they have same functional requirement and are of same type. In our implementation, we let communities perform tasks, and based on community task queue delay and also the assigned web service quality metric such as reliability, throughput, latency, successability and other metrics, the tasks will be performed with different qualities which we have evaluated and illustrated the results as average QoS of tasks performed}

%%%%%%%%%%%%%%%%%%%%%%%%% REVIEWER 3 %%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{2cm}

\newpage

\begin{center}
  \textbf{Reviewer 3}
\end{center}

\vspace{0.4cm}\textbf{\underline{1. Reviewer}:}  It is not convincing that the formation of service community can be modeled as a cooperative gaming problem. First, the concept of service community used in this paper is different from the ones used in the existing work, such as in the referred papers. The general idea of service community is a homogeneous service group that conceptually include services providing similar functionality. Services in the same community are independent and autonomous. There is no central control mechanism for the community. On the other hand, the concept of service community of this paper is a group of services which are monitored and managed by a community manager. It is not clear how realistic or practical such a community is and what the real world examples are.

\vspace{0.2cm}\textbf{\underline{Answer 1}: Most of the existing work such as [Liu 2012][18] (Coordinators), [Lim 2012][16] (Master Web Services], [Benatallah 2003][9] (Service Containers), [Maamar 2008][14], [Maamar 2009][8] (Master Web service), [Zeng 2003][11] (Service selection and task delegation), [Medjahed 2005][13] (Community Providers), [Limam 2010]\footnote{Managing Web Services Communities: A Cache for Queries Optimisation, International Journal on Web Service Computing, Vol.1, No.1, pp 2230-7702, 2010} (Centralized access in their community definition), have clearly stated that communities of web services need a centralized manager for the purposes of developing, coordinating and managing the community and in some of the works, being the access point of the community and having the task of delegating tasks.  We have added a web service community related work section to highlight this point.}

\vspace{0.4cm}\textbf{\underline{2. Reviewer}:}  The problem is modeled in a way so that the cooperative game theory can be applied, but not in a way that is consistent with the real world scenarios. The service parameters, such as its request rate and throughout rate are not in line with the general characteristics of web services, such as input and output parameters. It is not clear how the proposed service parameters are obtained and why they are important to determine a service's membership to a community, considering that they depend on not only the service's capacity, but also the amount and frequency of user requests, which can be random.

\vspace{0.2cm}\textbf{\underline{Answer 2}: For web services quality metrics, we have used a real world data set provided at: http://www.uoguelph.ca/\textasciitilde{}qmahmoud/qws/. Having a request rate for communities cannot be unrealistic, here is a real world example: Kooshiar Azimian. CEO of concertboom.com has provided me with their access and request rates of the web services they use\ref{cblog}. Their website provides ticketing information for events happening close to user's location. They use Ticketnetwork.com web services\footnote{Affiliate program link: http://www.ticketnetwork.com/affiliates/agreement.aspx} \footnote{Web Service API documentation: http://www.getacoder.com/data/projects/77613/TicketNetwork\%20Web\%20Services\%20(W2)-.doc}, \footnote{We were asked not to enclose these data publicly, thats why we could not add them to paper} and have implemented their back-end code in Java. They currently almost have 340,000 daily page views, so they call the ticketnetwork web service almost 4 times per second to get the seating chart and tickets available for the events. During a day, the pick time happens at around 20:00 UTC time, which has almost 10\% more request rates than around the lower pick which is around 7:00 AM UTC time. Therefore basically ticketnetwork, knows concertboom.com as a user has a request rate of 340,000 requests per day or almost 3.95 requests per second and they can adjust and change these thresholds, and adjust the internal service providers and resources they have to provide these rate of services.
}

\begin{figure}
\begin{center}
%\includegraphics[scale=0.6]{graph1Final+.eps}
\includegraphics[width=10cm]{cb.eps}\label{cblog}
\caption{Number of ticket pageviews of concertboom.com during May 5th, until Sep 9th, 2013}
\end{center}
\end{figure}

\vspace{0.4cm}\textbf{\underline{3. Reviewer}:}  The paper conducted a simulation instead of an experimental study using real world web services. This also suggests the difficulties of finding real world examples and application for the proposed idea.

\vspace{0.2cm}\textbf{\underline{Answer 3}: Our focus is on non-functional properties of web services and is extendable to any type of web service communities providing any type of service, which there is high demand, request load and competition for that service type such as mapping, weather, ticketing, and local places information. As mentioned in previous answer we used a real data set of web services (http://www.uoguelph.ca/\textasciitilde{}qmahmoud/qws/) for our implementations. We also added a functional SOAP (xml) based request, response to our implementation, however this as expected did not affect our simulation results, since our concern, as most of the related work, in on non-functional properties of web services. We added a modification to paper and highlighted this. In many related work such as
[Yu 2013]\footnote{Q. Yu and A. Bouguettaya, Efficient Service Skyline Computation for Composite Service Selection, IEEE Transactions on Knowledge and Data Engineering (TKDE), vo. 25, no. 4, pp. 776-789, 2013},
[Limam 2010]\footnote{Managing Web Services Communities: A Cache for Queries Optimisation, International Journal on Web Service Computing, Vol.1, No.1, pp 2230-7702, 2010}, and
[Liu 2012]\footnote{A. Liu, Q. Li, L. Huang, S. Ying, and M. Xiao,Coalitional game for community-based autonomous web services cooperation, IEEE Transactions on Services Computing, vol. 99, no. PrePrints, 2012.}
the results have been evaluated using simulations}

\vspace{0.4cm}

\vspace{1cm}

Finally, we hope that we have answered all the questions raised by
the reviewers, which will make the paper accepted for publication
in IEEE Transactions on Services Computing (TSC).

{
%\bibliographystyle{plain}
%\bibliography{omar}
}
\end{document}