### 2014-01-31 papers/coop-journal-R1/coop_response_letter.tex
changes

%\documentclass[preprint,authoryear,12pt]{elsarticle}
\documentclass[times, 12pt,a4paper]{article}
\usepackage{graphicx,amssymb,amsthm,stmaryrd}%,booktabs}
\usepackage{times}
\usepackage{amsmath,url,algorithm,algorithmic,mathrsfs,makeidx}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{epsfig}
\usepackage{latex8}

%journal{Journal of Knowledge-Based Systems}

\begin{document}
%\begin{frontmatter}

\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{example}{Example}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary} {Corollary}
\title{\underline{Response Letter}\\\vspace{0.5cm}To Compete or Cooperate? This is the Question in Communities of Autonomous Services}
\author{Ehsan Khosrowshahi Asl\\
e\_khosr@encs.concordia.ca\\
\and
Jamal Bentahar\\
bentahar@ciise.concordia.ca \\
\and
Rabeb Mizouni\\
rabeb.mizouni@kustar.ac.ae \\
\and
Babak Khosravifar\\
babak.khosravifar@mcgill.ca \\
\and
Hadi Otrok\\
hadi.otrok@kustar.ac.ae \\
}

\maketitle

First, we would like to thank and express our appreciation to the
reviewers for taking the time to carefully review our manuscript,
and for their insightful, valuable, and very useful comments. We
would like to thank the associate and chief editors as well for
their valuable comments and suggestions for improvements. We
carefully considered the comments and implemented all of them in
the revised version. In this letter, we will explain, point by
point, how the issues raised in the reviews are addressed and
answered. We hope our answers and revisions will satisfy the
reviewers and editors, which will make the paper accepted for
publication in Expert Systems With Applications Journal.\\

%\newpage

%%%%%%%%%%%%%%%%%%%%%%%%% REVIEWER 1 %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{center}
  \textbf{Reviewer 1}
\end{center}

\textbf{\underline{Reviewer}:} The presented paper is very good and up to date in the state of the art. The paper addresses a framework in which the agents use decision mechanisms in order to compete or cooperate.

\vspace{0.5cm} \textbf{\underline{1. Reviewer}:} The abstract should be shorter and it lacks information on the proposed framework and the decision mechanism based on game theory which are not mentioned. The suggested modifications aim make the abstract and the conclusion more similar.

\vspace{0.2cm}\textbf{\underline{Answer 1}: Thanks for the
comment. We made the abstract shorter and added the information about our decision making process which is based on ``game-theoretic best response technique'' to the abstract.}

\vspace{0.5cm}\textbf{\underline{2. Reviewer}:}  In section 2.2 (system parameters), the authors should specify how many parameters are used before explaining them and using such parameters in the proposed mathematics. This should aid the reader in better understanding.

\vspace{0.2cm}\textbf{\underline{Answer 2}: In the revised version, we added a short summary about the system parameters in section 2.2. We have five system parameters \emph{Task QoS}, \emph{Service QoS}, \emph{Budget}, \emph{Reputation} and \emph{Growth Factor}) and the other variables are derived or calculated based on these five parameters which are listed in Table 1. The title for Table 1. which was ``List of proposed system parameters'' was not very appropriate. We changed its title to ``List of abbreviations''}

\vspace{0.5cm}\textbf{\underline{3. Reviewer}:} The text consists on very long paragraphs, mainly section 4 (experimental results), which makes the reading difficult.

\vspace{0.2cm}\textbf{\underline{Answer 3}: We edited the text in the revised version.}

\vspace{0.5cm}\textbf{\underline{4. Reviewer}:} Still in section 4, the simulation application should be better specified, adding information on used machine, architecture, most important used services, etc.

\vspace{0.2cm}\textbf{\underline{Answer 4}: We agree. In the revised manuscript, we added more explanations and details about the architecture and service types. Each web service was modeled based on a class and we initiated many threads on a single process for the web services. They used abstract XML based messaging system with request parameters and a list of abstract XML based responses. Based on community task queue delay and the assigned web service quality metrics (such as reliability, throughput, latency and other metrics) these tasks will be performed with different qualities. In the experimental results and analysis section (Section 4), we have illustrated and
evaluated the average QoS of tasks performed. In the revised manuscript, we added more explanations and details about the architecture and service types.
}

\vspace{0.5cm}\textbf{\underline{5. Reviewer}:} Figure 3 is misplaced.

\vspace{0.2cm}\textbf{\underline{Answer 5}: Thanks for noticing the mistake, we replaced the Figure at the correct position (we put it after the paragraph it was first mentioned on) and double checked all other figures and tables placements.}

\vspace{0.5cm}\textbf{\underline{6. Reviewer}:} 6) There are some minor format errors that should be addressed.

\vspace{0.2cm}\textbf{\underline{Answer 6}: As mentioned in our Answer 4, we checked the formatting issues of paper as much as we could and we believe other formatting details will be covered the journal editors.
}

%\vspace{0.2cm}\textbf{\underline{Answer 9}: We
%\url{http://www.uoguelph.ca/\textasciitilde{}qmahmoud/qws/}, in}

%%%%%%%%%%%%%%%%%%%%%%%%% REVIEWER 2 %%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{2cm}

\newpage

\begin{center}
  \textbf{Reviewer 2}
\end{center}

\textbf{\underline{Reviewer}:} The paper perfectly fits within the scope of Expert Systems. This manuscript investigates the development of communities of autonomous services, formalizing the question "To compete or to cooperate?" between them. The authors do so by theorising and conducting a study through a dataset of 2,507 real services functioning on the web.

\vspace{0.5cm}\textbf{\underline{1. Reviewer}:}  Abstract: In my opinion, it is not necessary to talk about the sample used.

\vspace{0.2cm}\textbf{\underline{Answer 1}: True, as argued in our Answer 1 to Reviewer 1, We made the abstract more compact and removed some unnecessary details. }

\vspace{0.5cm} \textbf{\underline{2. Reviewer}}:  Introduction: The first phrase ``Services are software systems...'' should include which kind of services are being analysed: intelligent, autonomous,... The definition used does not correspond to ``Services'' as a whole.

\vspace{0.2cm}\textbf{\underline{Answer 2}: We agree, the definition was too broad we rephrased
the sentence as follows: Online Services are software systems providing a standard way of
interoperability among different applications. These services are developed
 in an autonomous and intelligent manner to continuously and autonomously interact with each other to
fulfill users' requests.}

\vspace{0.5cm} \textbf{\underline{3. Reviewer}}: Related work: The authors state in page 21: ``The interesting idea is to consider the benefits under four categories: innovation and learning,...''. If this is an interesting idea, it should be further explained. I agree with the authors, these benefits are very interesting, please provide further details.

\vspace{0.2cm}\textbf{\underline{Answer 3}: In the previous
version, the tasks were abstract. In the revised version, the
tasks are real tasks about flight booking that need to be.}

\vspace{0.5cm} \textbf{\underline{4. Reviewer}}:  Conclusion: The paper identifies clearly the implications for practice and further research. I consider the service consumer role very interesting for future research.

\vspace{0.2cm}\textbf{\underline{Answer 4}: We mean deviating from
the coalition. This has been fixed in the revised version.}

\end{document}