### 2014-03-04 seminar/seminar_report/chapter2.tex

\setcounter{chapter}{1}
%*******************************START Background **********************************

\chapter{Background and Relevant Literature}\label{sec:MAS}

In this chapter then we introduce
the concept of communities of web services, their architecture and
applications and the benefits of forming communities. Thereafter,
we discuss the cooperative game theory concepts used throughout
the proposal. Finally, we discuss relevant related work on web
service communities and games in the literature of service
oriented computing.

%\section{Community of Web Services}\label{sec:CommunityWS}
%In this section, we present web services and discuss the concept
%of their communities from architectural and operations
%perspectives.
%
%\subsection{Web Services}\label{sec:CWSWebServices}
%
%In our research work, we abstract web services as rational
%agents\footnote{The term
%        rational is used here in the sense that web services are utility
%        maximizers} providing services to end users. They aim to maximize
%        their individual income by receiving enough requests from end
%        users. In order to increase their revenue, web services seek for
%        more tasks if they have the capacity and throughput to do so. Web
%        services can join communities to have better efficiency by
%        collaborating with others, to have access to broad market share,
%        and to have opportunity of receiving a bigger task pool from end
%        users. Furthermore, the high reliance on web services has increased quality expectations from end users.
%        Communities of web services can provide higher availability, performance, reliability, and recovery for end users.
%
%\subsection{Web Service Communities}\label{sec:CWSDefinition}
% In \cite{DBLP:journals/internet/BenatallahSD03,
%Zeng:2003:QDW:775152.775211}, the authors introduce community of
%web services as collection of cooperative web services with common
%functionalities but different QoS metrics. Therefore
%$communities$ are differentiated from $composition$ types of web
%service cooperation in which web services with different
%functionalities work together to generate a new service with
%composite functionality.
%
%
%Maamar et al. in \cite{DBLP:journals/ijebr/MaamarSTBB09}
%proposed an architecture utilizing \emph{Contract-Net} protocol
%for engineering task distribution within communities. This
%architecture has been further developed in
%\cite{conf/IEEEscc/BenharrefSBB11, conf/IEEEscc/KhosravifarBMMT10,
%conf/aina/LimTM11, CSTintercommunity}. Two types of roles have
%been distinguished for community members: masters and slaves.
%Master web services lead communities and are responsible for
%membership management. They can invite and convince slave web
%services to join the community, and attract new slave web services
%to their communities by awarding them better payoff. Moreover,
%they can eject some slave members from the community to improve
%its overall reputation if these members are misbehaving or cannot
%provide the promised QoS \cite{DBLP:journals/ijebr/MaamarSTBB09}.

%        \begin{figure}
%            \begin{center}
%            \includegraphics[width=16cm]{Figures/wsarch.eps}\label{wsarch}
%            \caption{Communities of Web Services Architecture as Proposed in \cite{DBLP:journals/ijebr/MaamarSTBB09}}
%            \end{center}
%        \end{figure}
%
%Figure 1 %\ref{wsarch}
%depicts the basic architecture of
%communities of web services. The main components of the
%architecture are: 1) the providers of web services; 2) UDDI
%registries; and 3) communities platform. Communities abstract the
%same model of defining, announcing and invoking web services. They
%also adopt the same protocols that standard web services use with
%UDDI registries. UDDI is a platform-independent XML based registry
%list which facilitates worldwide web service discovery.

%The master web service is responsible with communication with web
%service providers and discovery registries. It it responsible for
%task distribution, web service selection, community management,
%maintaining a healthy set of web services satisfying end users
%requests with high QoS. In communities, the masters web services
%can be dedicated web services playing the master role during the
%entire time of being in the community. This master web service is
%independently developed and never participates in any
%composition. The master web services can also be chosen out of
%normal web services already inside the community
%\cite{DBLP:journals/ijebr/MaamarSTBB09}.

\section{Cooperative Game Theory and Multi-Agent Systems}\label{sec:CGTMS}

%        Cooperative game theory provides a set of mathematical and optimization tools for multi-agent environments. These tools have been utilized in communication networks and service oriented computing literature, where nodes as rational agents try to reason strategically and maximise their benefit.

Cooperative game is a branch of game theory that studies
strategies of self-interested entities or agents in a setting
where those agents can increase their payoff by binding agreements
and cooperating in groups. We let $N$ be a set of players which
can form a group called a $coalition$. A \emph{coalitional game}
is a pair $G = (N, v)$, where $v$ is called a \emph{characteristic
function} $v: 2^N \to \mathbb{R}$, mapping the set of players of
the coalition to a real number $v(N)$, the worth of $N$. This
number usually represents the output or payoff or again the
performance of these players working together as coalition. If a
coalition $S$ is formed, then it can divide its worth, $v(N)$ in
any possible way among its members. The payoff vector $x \in
\mathbb{R}^N$ is the amount of payoff being distributed among the
members of the coalition $N$.  A payoff vector is also \emph{efficient}
if the payoff obtained by a coalition is distributed amongst the
coalition members: $\sum_{i \in N} x_i = v(N)$.

        \subsection{Cooperative Game Concepts}

            {\bf Definition 1 (Shapley value)} Given a cooperative game $(N,
            v)$, the \emph{Shapley value} of player $i$ is given by \cite{shapley_value}:
            \begin{equation}\label{eq:shapley}
            \phi_i(N,v) = \sum_{S \subseteq N \backslash \left\{i\right\} }
            \frac{|S|! (|N|-|S|-1)!}{|N|!} (v(S \cup \left\{i\right\}) - v(S))
            \end{equation}

            \emph{Shapley value} is a unique and fair solution concept for
            payoff distribution among the members of the coalition.

            %\subsubsection{Core}

            {\bf Definition 2 (Core)} A payoff vector $x$ is in the $core$ of
            a coalitional game $(N, v)$ if and only if:
            \begin{equation}\label{eq:core}
            \forall S \subseteq N, \sum_{x_i \in S} x_i \geq v(S)
            \end{equation}

            The core is basically a set of payoff vectors where no subset of
            players $S^\prime$ could gain more than their current payoff by
            deviating and making their own coalition $\sum_{i \in S^\prime}
            x_i \geq v(S^\prime)$.

            {\bf Definition 3 (Convex cooperative games)} A game $(N,v)$ with
            characteristic function $v(S)$ is convex if:
            \begin{equation}\label{eq:convex}
            v(S) + v(T) \leq v(S \cup T) + v (S \cap T), \forall S,T \subseteq
            N.
            \end{equation}

            According to a classic result by Shapley \cite{S1971cores}, convex
            games always have a non-empty core. We will use a variation of
            convexity condition in our algorithm to check whether our
            coalitions are stable.

            %\subsubsection*{$\epsilon$-core}\label{s:epsilon}
%            %\emph{$\epsilon$-Core:}
%            %\\
%            When the \emph{core} set of a game is empty, it means no coalition
%            of players can gain anything by deviating. An outcome would be
%            unstable if a coalition can benefit even by a small amount from
%            deviating, which is a strong requirement. In fact, in some
%            situations, deviations can be costly, or players may have loyalty
%            to their coalitions, or even it can be computationally intractable
%            to find those small benefits. It would only make sense for a
%            coalition to deviate if the gain from a deviation exceeds the cost
%            of performing the deviation. \emph{$\epsilon$-core} relaxes the
%            notion of the core, and only requires that no coalition would
%            benefit significantly, or within a constant amount($\epsilon$) by
%            deviating (see Equation \ref{eq:core}).
%
%            \begin{equation}\label{eq:core2}
%            \forall S \subseteq N, \sum_{x_i \in S} x_i \geq v(S) - \epsilon
%            \end{equation}

%            \subsubsection*{Coalition Structure Formation}\label{sec:coalition}

            \emph{Coalition structure formation} is the problem of finding the best
            partition of web services into teams. In these settings, the
            performance of an individual service is less important than the
            \emph{social welfare} of the whole system, which is the sum of the
            values of all teams. Having the game $(N,v)$, a coalition
            structure $(CS)$ is \emph{socially optimal} if $CS$ belongs to set
            $\operatorname*{arg\,max}_{CS} v(CS)$ where $v(CS)$ is the sum of
            the values of all coalitions inside $CS$. $v(CS) = \sum_{C \in
            CS}v(C)$.
            %The outcome of a characteristic function game in coalition structure settings, consists of two parts; first a disjoint partition of players (agents) into coalitions, called a \emph{coalition structure} (CS) and second a \emph{payoff vector} as mentioned in cooperative game solution concepts, which distributes the value of each coalition among its members.

%        \subsection{Representation and Complexity Issues}\label{sec:CWSArchitecture}
%
%        Shapely value is the unique ``fair'' way to distribute the total surplus generated by the coalition, among all the players.
%        The nature of the Shapley value is combinatorial, as all possible orderings to form a
%        coalition needs to be considered. This computational complexity can sometimes be
%        an advantage as agents cannot benefit from manipulation. For example, it is NP-complete
%        to determine whether for a bunch of agents to collude and make their own coalition and guarantee
%        an increase in payoff of all participants \cite{conf/aaai/YokooCSOI05}.
%        There are some representations that allow us to compute the Shapley value efficiently by reducing the input size of the problem.
%        One example is \emph{Induced subgraph games}
%        which was introduced by Deng and Papadimitriou \cite{Deng94}. In this representation, players are represented by graph nodes, and
%        their valuation function should be the sum of weights of all edges between the node and all its neighbors. It is a succinct representation, using
%        an adjacency matrix, which needs only $O(n^2)$ space to store all the input, which is a major improvement from $O(2^n)$ because
%        if weights of all the edges in graph are all positive, the Shapley value can be computed in time $O(n^2)$.
%
%        In order to make cooperative game concepts practical in real world application, we have proposed an approximation multi-layer
%        algorithm useful for service oriented computing settings. Our excrements illustrate, these algorithms can provide
%        applicable and near optimal solutions for real world applications.

\subsection{Stability of Coalitions}

Core stability is a highly desirable property however in many problems it is not achievable. It would be more ideal to maintain a set of somewhat stable payoffs when the core is empty. There are several approaches to achieve this goal. One may drop the stability requirement and focus on other types of solutions for which a payoff division is guaranteed to exist. Two well known solution concepts in this category are \emph{nucleolus}\cite{schmeidler_nucleolus_1969} and the \emph{bargaining set}\cite{Davis67existenceof}. They try to minimize some measure of unhappiness in the game for the agents.

Another approach to stabilize the game can be achieved via external subsides. When \emph{Core} is empty it means the game is not stable since the coalition is unable to generate enough revenue to satisfy the demands of each subset of agents. An external party that is interested in stabilizing the game provides a subsidy to the agents if they form the grand coalition, and thus a value of $\lambda v(C)$ is divided among them, where $\lambda \geq 1$. Clearly any game can be stabilized using a large enough $\lambda$, however the external party would be interested in the minimal subsidy required in order to stabilize the game.

A community can also be stabilized by relaxing the core constraints. According to \emph{Core}, an outcome is unstable if a coalition can benefit even by a small amount from deviating, which is a strong requirement. In fact, in some situations, deviations can be costly, or players may have loyalty to their coalitions, or even it can be computationally intractable to find those small benefits. It would only make sense for a coalition to deviate if the gain from a deviation exceeds the cost of performing the deviation. \emph{$\epsilon$-core} relaxes the notion of the core, and only requires that no coalition would benefit significantly, or within a constant amount($\epsilon$) by deviating (see Equation \ref{eq:core}).

   \begin{equation}\label{eq:core2}
       \forall S \subseteq N, \sum_{x_i \in S} x_i \geq v(S) - \epsilon
   \end{equation}

Alternatively, the $\epsilon$ can be thought of as a \emph{tax} imposed on a coalition should it choose to deviate. This can again be seen as an external party, trying to stabilize the coalition by imposing some \emph{tax} on deviation. Taxation and subsiding as methods of stabilizing cooperative games have been studied in \cite{RePEc:spr:jogath:v:38:y:2009:i:1:p:3-16, Bachrach:2009:CSC:1692490.1692502, conf/ijcai/MeirRM11}.

%*******************************End Background **********************************