### 2014-03-04 seminar/seminar_report/chapter5.tex
changes

\setcounter{chapter}{3}
\chapter{Conclusion and Future Work}\label{sec:conclusionfuturework}

In this report, we proposed a cooperative game theory based model for the aggregation of web services within communities.
The goal of our services is to maximize efficiency by collaborating and forming stable
coalitions. Our method considers stability and fairness for all
web services within a community and offers an applicable mechanism
for membership requests and selection of web services. The
ultimate goal is to increase revenue by improving user
satisfaction, which comes from the ability to perform more tasks
with high quality. Simulation results show that our approximation
algorithms are polynomial in complexity and provide web services
and community owners with applicable and near-optimal decision
making mechanisms.

%As future work, we would like to perform more analytical and
%theoretical analysis on the convexity condition and also minimal $\epsilon$ values in \emph{$\epsilon$-core} solution concepts based on the
%characteristic function in web service applications. From web service perspective, the
%work can be extended to consider web service compositions where a
%group of web services having different set of skills cooperate to
%perform composite tasks. Also bargaining theory from cooperating
%game theory concepts can be used to help web services resolve the
%instability and unfairness issues by side payments.

\section {Future Plan and Timeline}

\indent The future goals in this Ph.D. research work are:

\begin{itemize}
\item \textbf{Winter 2014: }Analyzing other cooperative solution concepts such as Kernel
and Nucleolus where payoff division is guaranteed to exist and may
have optimal results.

\item \textbf{Summer 2014: }Analyzing the impact of Q-learning and reinforcement
learning on the performance of our model.

\item \textbf{Fall 2014: }Developing a community membership algorithm technique for
our agents in more distributed and ``incomplete information'' settings.

\item \textbf{Winter 2015: }Thesis work and PhD Defense.
%\item Developing an open source, Java based Tool, with UI for
%solving Core and Shapely solution concepts based on different
%input valuation functions.
\end{itemize}

%\begin{center}
%    \begin{tabular}{ | l | l | p{5cm} |}
%    \hline
%    Term & Activity & Publication \\ \hline
%    Winter 2014 & 1 & 1 \\ \hline
%    Summer 2014 & 1 & 1 \\ \hline
%    Fall 2014 & 1 & 1 \\ \hline
%    Winter 2015 & 1 & 1 \\
%    \hline
%    \end{tabular}
%\end{center}

\subsection {Publications }

    \begin{enumerate}
        \small{
        \item E. Khosrowshahi Asl, J. Bentahar, H. Otrok, R. Mizouni, \emph{``Efficient Community Formation for Web Services.''} Accepted at IEEE Transactions on Services Computing, Feb. 2014, IEEE. (Impact factor: 2.46)
        \item E. Khosrowshahi Asl, J. Bentahar, R. Mizouni, B. Khosravifar, H. Otrok, \emph{``To Compete or to Cooperate? This is the Question in Communities of Autonomous Services.''} In the Journal of Expert Systems and Applications, Feb. 2014, Elsevier. (Impact factor: 1.854, 5-year:  2.339)
        \item E. Khosrowshahi Asl, J. Bentahar, H. Otrok, R. Mizouni, \emph{``Efficient Coalition Formation for Web Services. In the 10th''} IEEE  International Conference on Services Computing (SCC 2013), Santa Clara Marriott, CA, USA, June 27-July 2, 2013
        \item B. Khosravifar, M. Alishahi, E. Khosrowshahi Asl, J. Bentahar, R. Mizouni, H. Otrok, \emph{``Analyzing Coopetition Strategies of Services within Communities''}. In the 10th International Conference on Service Oriented Computing, (ICSOC 2012), Shanghai, China, Nov. 2012
        }
    \end{enumerate}