### 2015-04-22 thesis/thesis/chap5/chap5.tex

\chapter{LEarning Chapter}\label{cha:PCTLKC+}

In this chapter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Summary}\label{sec:conclusion}

In this chapter,