### 2015-07-02 thesis/thesis/chap6/chap6.tex
changes

\chapter{Conclusion and Future Work}\label{Chap5:Conclusion}
%********************************************************************
%This chapter concludes the thesis. First, we give a summary of the main contributions of the thesis. Second, we present some hints for future directions.

%********************************************************************
\section{Conclusion}

In this theses we proposed three models for aggregation of web services within communities. The goal of our models are to maximize efficiency by collaborating and forming stable communities. In our first contributions focus on stability and

\section{Future Work}

Future Work Future Work Future Work

\newpage
\textbf{Publications in refereed journals and conferences}

\textbf{Journals}

\begin{itemize}
\item E. Khosrowshahi Asl, J. Bentahar, H. Otrok, R. Mizouni, "Efficient Coalition Formation for Web Services", IEEE Transactions on Services Computing, 2015.

\item E. Khosrowshahi Asl, J. Bentahar, H. Otrok, B. Khosravifar, R. Mizouni, "To compete or cooperate? This is the question in communities of autonomous services", Journal of Expert Systems with Applications, Elsevier, 2014.

\item F. Al-Saqqar, J. Bentahar, K. Sultan, W. Wan, E. Khosrowshahi Asl:, "Model checking temporal knowledge and commitments in multi-agent systems using reduction", Journal of Simulation Modelling Practice and Theory, 2015.

\end{itemize}

\textbf{Conferences}

\begin{itemize}
\item E. Khosrowshahi Asl, J. Bentahar, H. Otrok, R. Mizouni, "Efficient Community Formation for Web Services", IEEE SCC, Santa Clara, CA, USA, 2013.

\item B. Khoravifar, M. Alishahi, E. Khosrowshahi Asl, J. Bentahar, R. Mizouni, H. Otrok, "Analyzing Coopetition Strategies of Services within Communities", ICSOC, Shanghai, China, 2012

\item H. O. Marey, J. Bentahar, E. Khosrowshahi Asl, M. Mbarki, R. Dssouli, "Agents' Uncertainty in Argumentation-based Negotiation: Classification and Implementation", ANT/SEIT, Hasselt, Belgium, 2014.

\item H. Fallatah, J. Bentahar, E. Khosrowshahi Asl, "Social Network-Based Framework for Web Services Discovery", IEEE  FiCloud, Barcelona, Spain, 2014.

\end{itemize}

\textbf{Articles in process for publication in refereed journals}

\begin{itemize}
\item E. Khosrowshahi Asl, J. Bentahar, H. Otrok, R. Mizouni, "Distributed Decision Making for Dynamic Formation of Web Services Communities", Decision Support Systems, Elsevier (Submitted: June, 2015).
\end{itemize}