### 2015-07-16 thesis/thesis/ack.tex
changes

I would never have been able to finish my dissertation without the guidance of my committee members, help from friends, and support from my family.

I would like to thank my supervisors, Prof. Jamal Bentahar and Prof. Hadi Otrok for giving me the opportunity to work under their supervision. I am very grateful to them for their valuable suggestions and guidance throughout the preparation of this thesis. I learned a lot of valuable lessons which will be useful for me beyond the scope of this thesis throughout my lifetime. I also would like to thank Prof. Rabeb Mizouni for her significant help during my research work.

I would like to thank my examiner committee Professors Peter Grogono, Ferhat Khendek, Joey Paquet and Muhammad Younas for giving me the honor by being in my PhD committee. Their time and effort are greatly appreciated.

I would like thank my colleague Babak Khosravifar who was always willing to help and give his best suggestions. Also, I would like to thank my friends and lab colleagues Omar Marey, Faisal Al-Saqqar and Khalid Sultan for their help and support.

Finally, I am very grateful to my fiance and colleague Atieh Saberi and my parents for their understanding, encouragement and their endless support.